### 2022-12-20 paper/paper-mainfile.tex

%%%% ijcai22.tex

\typeout{IJCAI--22 Instructions for Authors}

% These are the instructions for authors for IJCAI-22.

\documentclass{article}
\pdfpagewidth=8.5in
\pdfpageheight=11in
% The file ijcai22.sty is NOT the same as previous years'
\usepackage{ijcai22}

% Use the postscript times font!
\usepackage{times}
\usepackage{soul}
\usepackage{url}
\usepackage[hidelinks]{hyperref}
\usepackage[utf8]{inputenc}
\usepackage[small]{caption}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{booktabs}
\usepackage{algorithm}
% \usepackage{algorithmic}
\urlstyle{same}

% the following package is optional:
%\usepackage{latexsym}

% See https://www.overleaf.com/learn/latex/theorems_and_proofs
% for a nice explanation of how to define new theorems, but keep
% in mind that the amsthm package is already included in this
% template and that you must *not* alter the styling.
\newtheorem{example}{Example}
\newtheorem{theorem}{Theorem}

% % % % % PACKAGES IMPORTED BY US % % % % %
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathbbol}
\usepackage[noend]{algpseudocode}
\usepackage{paralist}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{subcaption}
\usepackage{tikz}
\usetikzlibrary{aiplans}

% Following comment is from ijcai97-submit.tex:
% The preparation of these files was supported by Schlumberger Palo Alto
% Research, AT\&T Bell Laboratories, and Morgan Kaufmann Publishers.
% Shirley Jowell, of Morgan Kaufmann Publishers, and Peter F.
% Patel-Schneider, of AT\&T Bell Laboratories collaborated on their
% preparation.

% These instructions can be modified and used in other conferences as long
% as credit to the authors and supporting agencies is retained, this notice
% is not changed, and further modification or reuse is not restricted.
% Neither Shirley Jowell nor Peter F. Patel-Schneider can be listed as
% contacts for providing assistance without their prior permission.

% To use for other conferences, change references to files and the
% conference appropriate and use other authors, contacts, publishers, and
% organizations.
% Also change the deadline and address for returning papers and the length and
% page charge instructions.
% Put where the files are available in the appropriate places.

% PDF Info Is REQUIRED.
% Please **do not** include Title and Author information
\pdfinfo{
/TemplateVersion (IJCAI.2022.0)
}

\title{``Order is Half of Life'' -- Establishing Totally Ordered HTN Planning Problems is Half the Job (to Getting them Solved Efficiently)}

\author{Anonymous}

% Single author syntax
% \author{
%     Author Name
%     \affiliations
%     Affiliation
%     \emails
%     pcchair@ijcai-22.org
% }

% Multiple author syntax (remove the single-author syntax above and the \iffalse ... \fi here)
\iffalse
\author{
First Author$^1$
\and
Second Author$^2$\and
Third Author$^{2,3}$\And
Fourth Author$^4$
\affiliations
$^1$First Affiliation\\
$^2$Second Affiliation\\
$^3$Third Affiliation\\
$^4$Fourth Affiliation
\emails
\{first, second\}@example.com,
third@other.example.com,
fourth@example.com
}
\fi

\begin{document}

\maketitle

\begin{abstract}
HTN planning is in general undecidable, yet there exist efficient planning algorithms. An important subclass is that of totally ordered HTN planning problems. Not only is this class decidable, but also easier to handle algorithmically since tasks cannot interleave anymore. Consequently, several specialized solvers exist, and dedicated tracks at the International Planning Competition (IPC) in both 2020 and 2023. We were able to show that almost all currently existing partially ordered planning benchmarks can be turned into totally ordered ones without sacrificing solvability. Establishing totally ordered problems in a preprocessing technique thus allows to use planning systems, heuristics, pruning techniques (etc.) that are only available for totally ordered problems. This allows solving problems more efficiently and with better-informed techniques, at the (little) cost of doing a preprocessing technique. We show that in only a very few cases linearization isn't possible, and in all others establishing a total order pays off \emph{significantly}, thus establishing our technique as a significant ingredient in the state of the art for solving partially ordered HTN problems. As a side contribution, we identify a new class of decidable (EXPTIME-complete) HTN problems, namely those where our preprocessing technique is guaranteed to be solvability-preserving.
\end{abstract}

\input{00--macros}

\input{01--introduction}
\input{02--formalism}
\input{03--linearization-ground}
\input{04--otherLinearization}
\input{05--theory}
\input{06--evaluation}
\input{07--relatedWork}
\input{08--futureWork}
\input{09--conclusion}

%% The file named.bst is a bibliography style file for BibTeX 0.99c
\bibliographystyle{named}
\bibliography{bib}

\end{document}