### 2022-12-23 paper/table--POMethodBreak.tex

\begin{table}
	\caption{For the Problems in Each Domain, Average Percentage of Partially-Ordered Methods Linearized Without Cycle-break per Problem, for each Domain. Shows Results for Both Inference Methods.}\label{table:POMethodBreak}
	\begin{tabular}{lccl}
		\toprule
		Domain & Complex & Simple &  \\
		\midrule
		Transport & 1.0 & 0.0  \\
		UM-Translog & 0.8 & 0.0  \\
		Satellite & 1.0 & 0.0  \\
		Monroe Fully Observ. & 1.0 & 0.0  \\
		Rover & 1.0 & 0.0  \\
		Woodworking & 0.9 & 0.0  \\
		Monroe Part. Observ. & 1.0 & 0.0  \\
		PCP & 0.0 & 0.0  \\
		Barman-BDI & 1.0 & 0.0  \\
		SmartPhone & 0.1 & 0.1  \\
		Monroe & 0.8 & 0.0  \\
		Zenotravel & 1.0 & 0.0  \\
		\bottomrule
	\end{tabular}
\end{table}