### 2022-12-23 paper/table--SimpleLinearizable.tex

\begin{table}
	\centering
	\caption{These results are from using Simple Inference. Table shows the average number of methods per domain with more than 1 sub-task, the average number of those methods that needed cycle-breaking, and the percentage of successfully linearized methods.}
	\label{table:SimpleLinearizable}
		\scalebox{0.65} {
	\begin{tabular}{lcccl}
		\toprule
		& Avg. Num. Methods ($>$ 1) & Cycle-breaking & Percentage &  \\
		\midrule
Barman-BDI & 4870.0 & 3717.1 & 28.6  \\
Monroe & 12127.3 & 3447.8 & 72.4  \\
Monroe Fully Observ. & 49450.6 & 15813.0 & 68.0  \\
Monroe Part. Observ. & 49470.9 & 15809.3 & 68.0  \\
PCP & 12.3 & 12.3 & 0.0  \\
Rover & 509.4 & 355.2 & 32.3  \\
Satellite & 70.6 & 36.1 & 50.9  \\
SmartPhone & 6.3 & 4.0 & 21.8  \\
Transport & 1144.9 & 1144.9 & 0.0  \\
UM-Translog & 1.5 & 1.5 & 0.0  \\
Woodworking & 1927.3 & 1927.3 & 0.0  \\
Zenotravel & 190.2 & 99.4 & 47.0  \\
		\bottomrule
	\end{tabular}
}
\end{table}